---
title: "ActivityMonitor"
author: "Chunsu"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

# Reproducible Research: Course Project 1

```{r include=FALSE}
library(dplyr)
library(lubridate, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(magrittr, warn.conflicts = FALSE)
```


1. Read data
```{r}
    data <- read.csv("activity.csv", header = TRUE)
    data$date <- as.Date(data$date)
```

2. Histogram of the total number of steps taken each day

```{r, out.width = "55%", echo = TRUE}
##Calculate the total number of steps taken per day
    data %>% 
    group_by(date) %>%
    na.exclude %>%
    summarize(sum = sum(steps)) -> total_step
##Plot the histogram of the total number of steps taken each day
  	total_step %>% ggplot(aes(sum)) + 
    geom_histogram(fill = "steelblue", bins = 20, colour = "navy", alpha = 0.9) + 
    labs(title = "Histogram of Daily Total Number of Steps", 
       x = "Number of steps", 
       y = "Frequency") +
    theme_minimal()
```

3. Mean and median number of steps taken each day. Note how close the two values are.

```{r, out.width = "65%", echo = TRUE}

    central_values <- data.frame(measurement = c("mean", "median"), value = c(mean(total_step$sum), median(total_step$sum)))
    total_step %>% ggplot(aes(sum)) + 
    geom_histogram(fill = "steelblue", bins = 20, colour = "navy", alpha = 0.9) + 
    geom_vline(data = central_values, aes(xintercept = value, color = measurement), alpha =0.5) +
    scale_color_manual(values = c("red", "yellow"), name = NULL) +
    labs(title = "Histogram of Daily Total Number of Steps", 
       x = "Number of steps", 
       y = "Frequency") +
    theme_minimal()
    print(paste0("the mean number of steps is ", round(mean(total_step$sum),2)))
    print(paste0("the median number of steps is ", median(total_step$sum)))
```


4. Time series plot of the average number of steps taken
```{r, out.width = "65%", echo = TRUE}
    data %>% 
    group_by(interval) %>%
    na.exclude %>%
    summarise(mean_step = mean(steps)) -> imp_data 
    p <- ggplot(imp_data, aes(x=interval, y=mean_step)) +
    geom_line() + 
    xlab("")
    p
```

5. The 5-minute interval that, on average, contains the maximum number of steps
```{r, out.width = "65%", echo = TRUE}
  	p <- ggplot(imp_data, aes(x=interval, y=mean_step)) +
    geom_line() + 
    xlab("") +
    geom_vline(aes(xintercept = interval[which.max(mean_step)], color = "red")) +
    scale_color_manual(values = "red", name = NULL, labels = "max number of steps")
    p
    
```

6. Code to describe and show a strategy for imputing missing data
```{r}
    library(naniar)
```


First, we want to plot missing values column wise
```{r, out.width = "65%"}
    vis_miss(data)
```

```{r, out.width = "85%", echo = TRUE}
       ggplot(data,
       aes(x = interval, y = steps)) +
       geom_miss_point() + 
       facet_wrap(~date)
```

It seems that the missing data occurs on certain days, and there is a daily pattern throught out intervals (two peaks). So we imputate the missing data by calculating the interval avarage throught out the recording periods.
```{r}
  	data %>% 
    group_by(interval) %>%
    na.exclude %>%
    summarise(mean_step = mean(steps)) -> imp_data 
```

```{r}
##Collasping data across intervals within same date
    data %>% 
    group_by(date) %>%
    summarize(Mean = mean(steps)) -> avg_data
##Extracting a list of dates with NA
    na_date <- avg_data$date[which(is.na(avg_data$Mean), arr.ind=TRUE)]
##Creating new dataset with imputted values for NA
    new_data <- data
    for (i in 1: length(na_date)) {
      new_data$steps[new_data$date == na_date[i]] <- imp_data$mean_step
    }
    
```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r, out.width = "65%", echo = TRUE}

    new_data %>% 
    group_by(date) %>%
    na.exclude %>%
    summarize(sum = sum(steps)) -> total_step_new
      
    central_values <- data.frame(measurement = c("mean", "median"), value = c(mean(total_step_new$sum), median(total_step_new$sum)))
    total_step_new %>% ggplot(aes(sum)) + 
    geom_histogram(fill = "steelblue", bins = 20, colour = "navy", alpha = 0.9) + 
    geom_vline(data = central_values, aes(xintercept = value, color = measurement), alpha =0.5) +
    scale_color_manual(values = c("red", "yellow"), name = NULL) +
    labs(title = "Histogram of Daily Total Number of Steps", 
       x = "Number of steps", 
       y = "Frequency") +
    theme_minimal()
    print(paste0("the mean number of steps is ", round(mean(total_step_new$sum),2)))
    print(paste0("the median number of steps is ", round(median(total_step_new$sum),2)))
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r,out.width = "65%",echo = TRUE, message=FALSE}
      weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
      df <- mutate(data, weekday = weekdays(data$date))
      df$wDay <- factor((df$weekday %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') )
      ggplot(df, aes(x=interval, y=steps)) + 
        geom_point(alpha = 0.05) + 
        facet_grid(. ~ wDay) + 
        stat_summary(aes(y = steps,group=1,colour="red"), fun=mean, geom="line",group=1) +
        scale_color_discrete(labels = 'average') +
        theme(legend.title = element_blank())
```

