---
title: "ActivityMonitor"
author: "Chunsu"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research: Course Project 1

```{r include=FALSE}
library(dplyr)
library(lubridate, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(magrittr, warn.conflicts = FALSE)
```


1. Read data
```{r}
    data <- read.csv("activity.csv", header = TRUE)
    data$date <- as.Date(data$date)
```

2. Histogram of the total number of steps taken each day
```{r}
    data %>% 
    group_by(date) %>%
    na.exclude %>%
    summarize(sum = sum(steps)) %>%
  	with(hist(sum, xlab="total number of steps", main="Histogram of daily total number of steps"))
```

3. Mean and median number of steps taken each day
```{r}
  	data %>% 
    group_by(date) %>%
    na.exclude %>%
    summarize(Mean = mean(steps)) -> avg_data
```


```{r}    
    data %>% 
    group_by(date) %>%
    na.exclude %>%
    summarize(median = median(steps)) 
```

4. Time series plot of the average number of steps taken
```{r}
p <- ggplot(avg_data, aes(x=date, y=Mean)) +
  geom_line() + 
  xlab("")
p
```

5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
  	data %>% 
    group_by(interval) %>%
    na.exclude %>%
    summarize(mean_step = mean(steps)) %$%
    interval[which.max(mean_step)]
```

6. Code to describe and show a strategy for imputing missing data
```{r}
library(naniar)
```

```{r}
vis_miss(data)
```
First, we want to plot missing values
```{r}
       ggplot(data,
       aes(x = interval, y = steps)) +
       geom_miss_point() + 
       facet_wrap(~date)
```
It seems that the missing data occurs on certain days, and there is a daily pattern throught out intervals (two peaks). So we imputate the missing data by calculating the interval avarage throught out the recording periods.
```{r}
  	data %>% 
    group_by(interval) %>%
    na.exclude %>%
    summarise(mean_step = mean(steps)) -> imp_data 
```

```{r}
    data %>% 
    group_by(date) %>%
    summarize(Mean = mean(steps)) -> avg_data
    na_date <- avg_data$date[which(is.na(avg_data$Mean), arr.ind=TRUE)]
    for (i in 1: length(na_date)) {
      data$steps[data$date == na_date[i]] <- imp_data$mean_step
    }
    
```
7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
      data %>% 
      group_by(date) %>%
      summarize(Mean = mean(steps)) -> avg_data
      hist(avg_data$Mean)
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
      weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
      df <- mutate(data, weekday = weekdays(data$date))
      df$wDay <- factor((df$weekday %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') )
      ggplot(df, aes(x=interval, y=steps)) + 
        geom_point(alpha = 0.05) + 
        facet_grid(. ~ wDay) + 
        stat_summary(aes(y = steps,group=1,colour="red"), fun=mean, geom="line",group=1) +
        scale_color_discrete(labels = 'average') +
        theme(legend.title = element_blank())